---
title: "Tidyverse"
author: "Mitali Kulam"
format: "html"
editor: "visual"
---

## Tidyverse challange

Analyse the sale by location with a bar plot. since state and cities are multiple features,they should be split.Replace your bike_orderlines_wrangled_tbl object with the newly wrangled object.

##Running Code

#2.0 Load required libraries:

```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)
library(writexl)
```

##2.1 Read the data from the files, drive it from the given path

```{r}
bikes_tbl <- read_excel(path = "C:/Users/dell/Desktop/business data science/tidyverse/ss23-bdsb-mitalikulam/assets/tidyverse/bikes.xlsx")
orderlines_tbl <- read_excel(path = "C:/Users/dell/Desktop/business data science/tidyverse/ss23-bdsb-mitalikulam/assets/tidyverse/orderlines.xlsx")
bikeshops_tbl  <- read_excel(path = "C:/Users/dell/Desktop/business data science/tidyverse/ss23-bdsb-mitalikulam/assets/tidyverse/bikeshops.xlsx")
```

### 2.2 Examin Data

```{r}
#print it to the console
orderlines_tbl

```

### 2.3 Joining Data

```{r}
left_join(orderlines_tbl,bikes_tbl, c("product.id" = "bike.id"))

bike_orderlines_joined_tbl <-orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))

bike_orderlines_joined_tbl
```

### 2.4 manipulate the data

```{r}
bike_states_wrangled_tbl <- bike_orderlines_joined_tbl %>%
  # 5.1 Separate state and city name
  separate(location, into = c("city", "state"), sep = ", ") %>%
  

 
  # 5.2 Add the total price (price * quantity) 
  # Add a column to a tibble that uses a formula-style calculation of other columns
  mutate(total.price = price * quantity)
```

##### Select columns and add a year followed by summerizing year and state and then format

```{r}
sales_by_states_tbl <- bike_states_wrangled_tbl %>%



  select(order.date, total.price, state) %>%
  mutate(year = year(order.date)) %>%
  
  # Group by and summarize year and states
  group_by(year, state) %>%
  summarise(sales = sum(total.price)) %>%
  ungroup() %>%
  
  
  # Format $ Text
  mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))

sales_by_states_tbl 
```

## visulation by getting a graph

```{r}
sales_by_states_tbl %>%
  
  # Set up x, y, fill
  ggplot(aes(x = year, y = sales, fill = state)) +
  
  # to your plotting code to rotate your x-axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  
  # Geometries
  geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
  geom_label(aes(label = sales)) + # Adding labels to the bars
  geom_smooth(method = "lm", se = FALSE) + # Adding a trendline
  
  # Facet
  facet_wrap(~ state) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title = "Revenue by state",
    subtitle = "Each state has an upward trend",
    fill = "States" # Changes the legend name
  )
```
