---
title: "Data Acquisition"
author: "Mitali Kulam"
format: "html"
editor: "visual"
---

# Load libraries

```{r}
library(RSQLite)
library(dplyr)
library(httr)
library(glue)
library(jsonlite)
library(keyring)
library("rstudioapi")
library(rvest)

```

# -- Dtabase Connection object

## Challenge Task

# -- 1) Get some data via an API. There are millions of providers, that offer API access for free and have good documentation about how to query their service.

# Send the GET request to the API

```{r}

response <- GET("https://www.themealdb.com/api/json/v1/1/filter.php?c=Seafood")

```

# Check the status code

```{r}

if (response$status_code == 200) {
  # The request was successful
  # Get the response data
  rawToChar(response$content)
  
  meals_db <- response %>%
    .$content %>%
    rawToChar() %>%
    fromJSON()
  #-- Print the first 10 rows of your tibbles
  head(meals_db)
} else {
  # The request failed
  print(response$status_code)
}
```

# -- 2) Scrape one of the competitor websites of canyon

# Set the URL of the website to scrape

```{r}
url <- "https://www.rosebikes.de/"

```

# Scrape the website

```{r}
html <- read_html(url)
```

# Find the elements that contain the model names

```{r}

model_names <- html %>%
  html_nodes(".products-slider-tile__product-name") %>%
  html_text()
```

# Find the elements that contain the prices

```{r}
prices <- html %>%
  html_nodes(".product-tile-price__current-value") %>%
  html_text()
```

# Extract the number using regular expressions

```{r}

price <- gsub("[^0-9.,]", "", prices)  # Remove all non-digit, non-decimal, and non-comma characters

```

# Create a data frame of the model names and prices

```{r}
database <- data.frame(
  Model = model_names,
  Price = price
)
```

# Print the data frame

```{r}
print(database)
```
