---
title: "Data Wrangling"
author: "Mitali Kulam"
format: "html"
editor: "visual"
---

##2.0 Running code ###2.1 Load Libraries

```{r}
library(tidyverse)
library(vroom)
library(data.table)
library(tictoc)
library(dplyr)
library(readxl)

```

### 2.2 Data import

## Define column type for vroom

```{r}
col_types <- list(
  id = col_character(),
  type = col_character(),
  number = col_character(),
  country = col_character(),
  date = col_date("%Y-%m-%d"),
  abstract = col_character(),
  title = col_character(),
  kind = col_character(),
  num_claims = col_double(),
  filename = col_character(),
  withdrawn = col_double()
)
```

# -- Using Alternative 2: The reduced data set for the year 2014

## Read the data into a tibble

```{r}
patent_tbl <- vroom(
  file = "C:/Users/dell/Desktop/business data science/tidyverse/ss23-bdsb-mitalikulam/assets/wrangling/patent.tsv", 
  delim = "\t"
  #col_types = col_types,
  #na = c("", "NA", "NULL")
)

assignee_tbl <- vroom(
  file = "C:/Users/dell/Desktop/business data science/tidyverse/ss23-bdsb-mitalikulam/assets/wrangling/assignee.tsv", 
  delim = "\t", 
  col_types = col_types,
  na = c("", "NA", "NULL")
)

patent_assignee_tbl <- vroom(
  file = "C:/Users/dell/Desktop/business data science/tidyverse/ss23-bdsb-mitalikulam/assets/wrangling/patent_assignee.tsv", 
  delim = "\t", 
  col_types = col_types,
  na = c("", "NA", "NULL")
)

uspc_tbl <- vroom(
  file = "C:/Users/dell/Desktop/business data science/tidyverse/ss23-bdsb-mitalikulam/assets/wrangling/uspc.tsv", 
  delim = "\t", 
  col_types = col_types,
  na = c("", "NA", "NULL")
)
```

## Join Tables

```{r}
patent_joined_table <- patent_assignee_tbl %>%
  inner_join(assignee_tbl, by = c("assignee_id" = "id")) %>%
  inner_join(patent_tbl, by = c("patent_id" = "id"))

```

# Patent Dominance: What US company/corporation has the most patents?

## List the 10 US companies with the most assigned/granted patents.

```{r}
us_patent_counts <- patent_joined_table %>%
  group_by(organization) %>%
  summarise(total_patents = n()) %>%
  arrange(desc(total_patents))

top_10_us_companies <- head(us_patent_counts, 10)

top_10_us_companies
```

##What US organization had the most patents granted in August 2014? List the top 10 companies with the most new granted patents for August 2014.

```{r}
us_patent_counts_aug_2014 <- patent_joined_table %>%
  filter(format(date, "%Y-%m") == "2014-08") %>%
  group_by(organization) %>%
  summarise(total_patents = n()) %>%
  arrange(desc(total_patents)) %>%
  select(organization, total_patents)

top_10_us_companies_aug_2014 <- head(us_patent_counts_aug_2014, 10)

top_10_us_companies_aug_2014
```
